\chapter{Machine Learning in the industry}
%\newrefsection

\section{Real-world ML Challenges at the Scale of Banking by BBVA Data and
Analytics}

Up to this moment, ML in banking has focused on fraud detection.

Working at the moment on an app for expenses prediction. Hundreds of millions
of time series.

Classification of transactions, they tested Word2vec with Vector of Locally
Aggregated Descriptors (VLAD) pooling.

Expense forecasting with LSTM layers and a dense layer on top. Now the expenses
forecasting is including uncertainty on the predictions (See Uncertainty
  Modelling in Deep Networks: Forecasting Short and Noisy Series
\cite{brando2018uncertainty})

How to compare customers (see cleint2vec: Towards systematic baselines for
banking applications \cite{baldassini2018client2vec})

Recommender systems, modelling users for target advertising.

In order to create a loss functions that considers (See, A Missing Information
Loss function for implicit feedback datasets \cite{arevalo2018missing} )

Reinforcement Learning for Fair Dynamic Pricing \cite{maestre2018reinforcement}

What is not in the books of Machine Learning when applying methods in the
industry:

\begin{itemize}
  \item Fairness
  \item Privacy
  \item Ethics
  \item Business
  \item Data acquisition/labelling
  \item UX/UI
  \item Design
\end{itemize}

\section{Data Efficient Reinforcement Learning by PROWLER.io}

Based in three teams: Probabilistic team, reinforcement learning team, and multiagent team.

Currne tReinforcement learning technology, successes using deep Q-Networks.

\section{Lynx: real-time accurate fraud detection over massive data.
Instituto de Ingenieria del conocimento by Álvaro Barbero Jiménez}

There are 8 countries using Lynx, and over 30,000 million transactions
processed per year.

Their system need to make decision in a few miliseconds.

Device tries to make an operation, this is sent to the institution (bank), the
institution sends a copy of the operation to Lynx, that will estimate the
probability of fault in some miliseconds and send the answer back to the
institution. Then the institution has to choose on what action to take and send
back the operation success (or deny) to the device.

The decisions are taken with two types of analysis

\begin{enumerate}
  \item Parametric Analysis
  \item \dots
\end{enumerate}

Approach for programming: Use Python, R, elastic and cocker in order to test
ideas fast, and if the results are good, implement it on C, Bash (see book Data
science at the command line) or Fortran.

One of the problems of detecting fraud is that the fraudulent transactions
evolve as they are detected. It is necessary to use Adaptive models and train
with incremental learning.

Hardware specification: 384 threads, 6TB RAM, 40 TB SSD (a training in one day)
Training data 800k-5M transactions.

As an error measure they use Value Detection Rate.

\section{Microsoft Research by Sebastian Nowozin}

120 researchers at Cambridge (200 total workers). Cambridge focus on Machine
Learning for healthcare. New lab in Montreal focused on Reinforcement learning
and Deep Learning.

Why is Artificial Intelligence growing? (1) Massive computation power (GPUs,
FPGAs, \dots), (2) Powerful algorithms (2012 Alexnet?), (3) Big data.

AI principles

\begin{itemize}
  \item Fairness: we want the algorithms to avoid systematic biases. It is
    difficult to remove biases that are already in the datasets.
  \item Accountability:
  \item Transparency:
  \item Ethics:
\end{itemize}

\begin{mybox}
  If you know what you are doing, then you are not doing research. (not sure if
  this was the exact wording) -- Sebastian Nowozin
\end{mybox}

\subsection{Timelines}

In images, videos and audio: 2000 basic research in audio, skeletal tracking
and facial recognition, 2010 Kinect, 2011 Kinect Fusion, 2012 HoloDesk, 2015
HoloLens.

In translation: 1991 basic research in natural language and speech recognition,
2007 Big bets with product team, 2014 promise of speech recognition with
translation, 2015 Skype translator launches, 2016 Microsoft translator API and
personal universal translator launches.

Visual Studio IntelliCode (AI-assisted development)

InnerEye, how can computers understand the segments of medical images.
Previously, a doctor had to spend around 8 hours segmenting a tumour in several
images. Now, a program can do the segmentation really fast, and the doctor can
check the results much faster.

Adaptive, learning to decode the immune system to diagnose disease. From a
blood sample, immunosequencing extracts some features from t-cells?, a Machine
learning model can be trained in order to improve the health care service.
